---
title: "Balance Analysis"
author: "Susan Glenn, Jeanne Li, Jingcheng Xu, Stewart Kerr"
date: "12/2/2019"
output: pdf_document
---

In terms of balance, the matching improved balance significantly. All of our after matching standardized differences were below 0.1. Note that we only checked balance on baseline covariates and threw our all of the matched pairs where subjects died in the same wave or lived throughout the followup period. Another possible avenue to address balance issues is enforcing the balance constraints in our risk set matching. This would force balance among important covariates. Alternatively, we could also do risk set matching allowing for matching with multiple controls. Out of the 16 baseline covariates that balance was checked on seven that had a higher standardized differences before matching.The code to assess balance is below.

### Balance
```{r}
library(dplyr)
library(ggplot2)
library(compare)
#Load matches
after <- read.csv("../data/results-final.csv")
#Load data before matching
before <- read.csv("../data/data-stacked.csv")
#Create a vector of columns for which we want to check balance
balance_cols = colnames(after)[9:ncol(after)]

#making an outcome variable for before matching
before$treated[before$FIRST_WS == -1] = 0
before$treated[before$FIRST_WS > 0] = 1

#Before matching
treatedmat = subset(before, treated==1)
controlmat = subset(before, treated==0)

#Standardized differences before matching
treatedmat.X.before <- treatedmat[,balance_cols]
controlmat.X.before <- controlmat[,balance_cols]
controlmean.before = apply(controlmat.X.before, 2, mean, na.rm=TRUE);
treatmean.before = apply(treatedmat.X.before, 2, mean, na.rm=TRUE);
treatvar = apply(treatedmat.X.before, 2, var, na.rm=TRUE);
controlvar = apply(controlmat.X.before, 2, var, na.rm=TRUE);
stand.diff.before = (treatmean.before-controlmean.before)/sqrt((treatvar+controlvar)/2)

##After matching
controlmat = subset(after, treated==0)
treatedmat = subset(after, treated==1)

#get rid of outcome, treated, id variables etc..
controlmat.X.after <- controlmat[,balance_cols]
treatedmat.X.after <- treatedmat[,balance_cols]
controlmean.after = apply(controlmat.X.after, 2, mean)
treatedmean.after = apply(treatedmat.X.after, 2, mean)
#treatvar.after = apply(treatedmat.X.after, 2, var, na.rm=TRUE);
#controlvar.after = apply(controlmat.X.after, 2, var, na.rm=TRUE);

# Standardized differences after matching
stand.diff.after = (treatedmean.after-controlmean.after)/sqrt((treatvar+controlvar)/2);

##Making table standardized differences
standBeforeAfter = cbind(stand.diff.before, stand.diff.after)
colnames(standBeforeAfter ) = c("Before Match (Standardized Diff)",
                                "After Match (Standardized Diff)")
knitr::kable(round(abs(standBeforeAfter),3), caption = "Differences in Covariates (Before and After)") 

### Love Plot
abs.stand.diff.before=abs(stand.diff.before)
abs.stand.diff.after=abs(stand.diff.after)
covariates=names(stand.diff.before)

plot.dataframe = data.frame(abs.stand.diff=c(abs.stand.diff.before,abs.stand.diff.after), covariates=rep(covariates,2), type=c(rep("Before",length(covariates)), rep("After",length(covariates))))

p1 <- ggplot(plot.dataframe,aes(x=abs.stand.diff,y=covariates))+geom_point(size=5,aes(shape=factor(type)))+scale_shape_manual(values=c(4,1))+geom_vline(xintercept=c(.1,.2),lty=2)
```


###Balance after CART
```{r}
after_match <- read.csv("../data/results-final.csv")
#Load matches
matched_pairs.CART <- read.csv("../data/matched-pairs.CART.csv")
after_CART <- subset(after_match, pair_ID %in% matched_pairs.CART$pair_ID)

##After matching
controlmat = subset(after_match, treated==0)
treatedmat = subset(after_match, treated==1)

#get rid of outcome, treated, id variables etc..
x <- intersect(names(stand.diff.before), names(controlmat))
controlmat.after <- controlmat %>%
  dplyr::select(x)
treatedmat.after <- treatedmat %>%
  dplyr::select(x)

controlmean.after = apply(controlmat.after, 2, mean)
treatedmean.after = apply(treatedmat.after, 2, mean)
#treatvar = apply(treatedmat.after, 2, var, na.rm=TRUE);
#controlvar = apply(controlmat.after, 2, var, na.rm=TRUE);

# Standardized differences after matching
stand.diff.after = (treatedmean.after-controlmean.after)/sqrt((treatvar+controlvar)/2);

##After CART
controlmat = subset(after_CART, treated==0)
treatedmat = subset(after_CART, treated==1)
#get rid of outcome, treated, id variables etc..
x <- intersect(names(stand.diff.before), names(controlmat))

controlmat.after <- controlmat %>%
  dplyr::select(x)
treatedmat.after <- treatedmat %>%
  dplyr::select(x)

controlmean.after = apply(controlmat.after, 2, mean)
treatedmean.after = apply(treatedmat.after, 2, mean)
#treatvar = apply(treatedmat.after, 2, var, na.rm=TRUE);
#controlvar = apply(controlmat.after, 2, var, na.rm=TRUE);
# Standardized differences after matching
stand.diff.after.CART = (treatedmean.after-controlmean.after)/sqrt((treatvar+controlvar)/2);
##Making table standardized differences
standBeforeAfter = cbind(stand.diff.before, stand.diff.after, stand.diff.after.CART)
colnames(standBeforeAfter ) = c("Before Match (Standardized Diff)",
                                "After Match (Standardized Diff)",
                                "After CART (Standardized Diff)")
knitr::kable(round(abs(standBeforeAfter),3), caption = "Differences in Covariates (Before and After)") 



### Love Plot
abs.stand.diff.before <- abs(stand.diff.before)
abs.stand.diff.after <- abs(stand.diff.after)
abs.stand.diff.after.C <- abs(stand.diff.after.CART)
covariates <- names(stand.diff.before)

plot.dataframe <- data.frame(abs.stand.diff=c(abs.stand.diff.before,abs.stand.diff.after,abs.stand.diff.after.C), covariates=rep(covariates,3), type = c(rep("Before",length(covariates)), rep("After Matching",length(covariates)), rep("After CART",length(covariates))))

p2 <- ggplot(plot.dataframe, aes(x = abs.stand.diff, y = covariates, color=type)) + 
  geom_point(size=5, aes(shape = factor(type))) +
  scale_shape_manual(values = c(4,1,6)) + 
  geom_vline(xintercept=c(.1,.2), lty=3)
```